\vspace{-15mm}\fontsize{16pt}{10pt}\selectfont\textbf{Facial Expression Recognition with\\ Convolutional Neural Networks}

### 2018-04-30 doc/paper.tex
move figure

\subsection{Training Procedure}
The CNN is trained on a subset of the FER-dataset according to the original train/test/validation split with the FERplus prabability distributions as labels. The models filters and weights/biases of the fully connected layers are fit to the training set using a backpropagation algorithm. Therefore, a cross-entropy loss function is used to measure the model error and the Adam optimization algorithm with an initial step size of 0.001 is used to minimize the loss. To prevent overfitting the model on the training set, the loss value on the validation set (unseen data) is computed periodically during training. When the validation loss stops decreasing, training is stopped and the final model accuracy is computed on the test subset.

@@

% - CNN robustness to occlusion, distortions and geometric transformations.
The final model scored an accuracy of ~75\% on a test-set of 3557 samples. In live predictions with a webcam, the classifier has shown to be accurate. However, predictions for the classes $disgust$ and $fear$ are hard to reproduce. This is most likely because the seven classes of emotion in the dataset are highly imbalanced. For instance disgust only has 500 samples while happy has much more. This results in biased predictions with higher probabilities of the majority classes. In order to balance the training dataset, data augmentation could be used \cite{krizhevsky12}: To increase the number of training examples for a given class, the images could be flipped horizontally or rotated in multiple random angles. Given more time for this project, I would like to have experimented with such techniques of data augmentation.

>>>>

\subsection{Training Procedure}
The CNN is trained on a subset of the FER-dataset according to the original train-test-validation split with the FERplus prabability distributions as labels. The models filters and weights/biases of the fully connected layers are fit to the training set using a backpropagation algorithm. Therefore, a cross-entropy loss function is used to measure the model error and the Adam optimization algorithm with an initial step size of 0.001 is used to minimize the loss. To prevent overfitting the model on the training set, the loss value on the validation set (unseen data) is computed periodically during training. When the validation loss stops decreasing, training is stopped and the final model accuracy is computed on the test subset.

@@

% - CNN robustness to occlusion, distortions and geometric transformations.
The final model scored an accuracy of ~75\% on a test-set of 3557 samples. In live predictions with a webcam, the classifier has shown to be accurate. However, predictions for the classes $disgust$ and $fear$ are hard to reproduce. This is most likely because the seven classes of emotion in the dataset are highly imbalanced. For instance disgust only has 500 samples while happy has much more. This results in biased predictions with higher probabilities of the majority classes. In order to balance the training dataset, data augmentation could be used \cite{krizhevsky12}: To increase the number of training examples for a given class, the images could be flipped horizontally or rotated in multiple random angles. Given more time for this project, I would like to have experimented with such techniques of data augmentation and also with  more complex model architectures.
